---
title: "Project 2"
author: "Lin Ye"
date: "October 6, 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(tidyr)
library(stringr)



```
In This project, Im going to analyze the store performance. The tables are coming in messy format, and I will transform the tables to achieve my analysis

##Finding Total Sales, and Percentage of Two Regions

The original table listed East and West in two seperate colums.

Getting data, display messy form
```{r}
FilePath <-"https://raw.githubusercontent.com/yelin0000/607HW/master/sales_summary.csv";
sales_temp <-read.csv(FilePath, sep=",")

head(sales_temp)
dim(sales_temp)

```
Start transformation
```{r}

sales_temp1 <- gather(sales_temp,key,Value,-Day, -WeekDay:-DayOfYear, na.rm = TRUE)
sales_temp1
```

Once we tidy up the data we calculate the total sales and percentage of each region. From the data, West is generating slight more sales than East
```{r}

sales_temp2 <- sales_temp1 %>% 
  separate(key, c("Region", "Metrics"), 4) %>% 
  spread(Metrics, Value)  %>%
  arrange(Region,DayOfYear)
sales_temp2

TotalSales<-sum(sales_temp2$NetSales)

EastPercent<-sum(sales_temp2$NetSales[sales_temp2$Region=="East"])*100/TotalSales
WestPercent<-sum(sales_temp2$NetSales[sales_temp2$Region=="West"])*100/TotalSales

c(TotalSales, EastPercent,WestPercent)

```



##Sales Qty by Week

In this sample, the table listed units sold in 4 weeks of Oct. We want to list the units sold by week

Getting data 
```{r}
FilePath2 <-"https://raw.githubusercontent.com/yelin0000/607HW/master/Sales.csv";
sales_temp3 <-read.csv(FilePath2, sep=",")

sales_temp3
```
Tidy data
```{r}

sales_temp4 <- gather(sales_temp3,key,Qty,-Account.Number:-Week, na.rm = TRUE) 
tail(sales_temp4)


sales_temp5 <- sales_temp4 %>% 
  filter(Qty !="-") %>%
  separate(key, c("X","WeekBegin"),1) %>%
  select (-c(Week,X)) 


sales_temp6<-mutate(sales_temp5,StoreName = ifelse(grepl("919", Account.Number), "Flatiron","BroadWay"))

sales_temp6


```

##Average Sales per Transactions
In this example, we want to find out the spread of change of average sales per trasaction year over year by day of week 

```{r}

FilePath3 <-"https://raw.githubusercontent.com/yelin0000/607HW/master/AST.csv";
  
#"C:\\Users\\lin\\Documents\\DB\\AST.csv";

AST_temp<-read.csv(FilePath3,sep=",")
AST_temp
```

tidy data.

The change of AST year over year is on average 10%
```{r}
AST_temp1<-mutate(AST_temp,YearLabel = ifelse(grepl("2017", Year), "ThisYear","LastYear")) 

AST_temp2<-  gather(AST_temp1,Month,AST,-Year,-YearLabel,-WeekDay) %>%
  arrange(Month,WeekDay) %>%
  select(-c(Year)) %>%
  spread(YearLabel, AST) 

ASt_temp3 <- mutate(AST_temp2,AST_Diff=ThisYear-LastYear,AST_Chg=((ThisYear/LastYear)-1)*100)

summary(ASt_temp3$AST_Chg)


```